---
output: github_document
---

# hfaggregate

`hfaggregate` provides utilities for collapsing high-resolution hydrofabrics
into analysis-ready outlet networks or uniform-sized catchment distributions.
It builds on the [`hfutils`](https://github.com/lynker-spatial/hfutils) toolkit
and expects standard Hydrofabric GeoPackages as inputs.

## Installation

```r
# install.packages("remotes")
remotes::install_github("lynker-spatial/hfaggregate", dependencies = TRUE)
```

The package imports `sf`, `terra`, `dplyr`, `hfutils`, and `cli`; make sure the
system requirements for those packages (GEOS, GDAL, PROJ) are available.

## Key functionality

- `aggregate_to_outlets()`: unions divides and mainstem flowpaths by outlet
  groups discovered from POIs or critical levelpath junctions.
- `aggregate_to_distribution()`: enforces target catchment sizes while
  respecting minimum flowpath length/area constraints and optional POI locks.

## Quick start

```{r, warning=FALSE, message=FALSE}
library(hfutils)
library(hfaggregate)

hf_path <- "/Users/mikejohnson/hydrofabric/hydrofabric-v3/tmp/rfc/nwis-06752260_rfc.gpkg"

pois  <- hfutils::as_ogr(hf_path, "pois") |> 
  dplyr::select(flowpath_id, poi_id, geom) |> 
  hfutils::st_as_sf()

# Collapse the network around POI-constrained outlets
outlet_net <- aggregate_to_outlets(
  gpkg     = hf_path,
  pois     = pois
)

hfsubset::hfview(outlet_net) + pois
```

```{r, warning=FALSE, message=FALSE}
# Build a uniformly sized distribution
distribution <- aggregate_to_distribution(
  gpkg = hf_path,
  pois = pois
)

hfsubset::hfview(distribution) + pois
```

Both functions return lists with `flowpaths`, `divides`, and (optionally) `pois`
layers that can be written to disk via `hfutils::write_hydrofabric()`.

## Tuning aggregation behaviour

- **Outlets vs. POIs**: `aggregate_to_outlets()` seeds groups from the supplied
  POIs; set `auto_seed_terminals = TRUE` inside `.partition_and_flag_mainstem()`
  if you need terminal nodes treated as outlets even without POIs.
- **Size constraints**: `aggregate_to_distribution()` first merges along
  mainstems (`aggregate_along_mainstems()`), then collapses headwaters via
  `collapse_headwaters()`. Adjust `ideal_size_sqkm`, `min_length_km`, and
  `min_area_sqkm` to bias the result toward longer or wider units.
- **POI preservation**: Provide an `sf` data frame with `flowpath_id` and
  `poi_id` columns to prevent POI-bearing segments from being merged away.


